---
title: "432 data clean"
output: pdf_document
date: "2023-05-07"
---

```{r}
df <- read.csv("home_insurance.csv", header=T, na.strings=c("","NA"))
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

POL_STATUS is the response variable of our classification problem. We first delete all missing POL_STATUS.

```{r}
na_counts <- colSums(is.na(df))
print(na_counts)
na_cols <- names(na_counts[na_counts > 30000])
na_cols
```
Delete columns with too many missing values and inmeaningful
```{r}
columns_to_exclude <- c("P1_PT_EMP_STATUS", "CLERICAL", "CAMPAIGN_DESC", "UNSPEC_HRP_PREM", "Police",
                        "P1_SEX", "SUBSIDENCE", "HP1_ADDON_PRE_REN", "HP1_ADDON_POST_REN",
                        "HP2_ADDON_PRE_REN", "HP2_ADDON_POST_REN", "HP3_ADDON_PRE_REN", 
                        "HP3_ADDON_POST_REN",
                        "GARDEN_ADDON_PRE_REN", "GARDEN_ADDON_POST_REN", "i","QUOTE_DATE")
df <- df[, !names(df) %in% columns_to_exclude]
```

POL_STATUS is the response variable of our classification problem. Our response variable MTA_APRP for regression problem. We drop all the missing value of these two variables.
```{r}
df<-df %>% drop_na(POL_STATUS)
df<-df %>% drop_na(MTA_APRP)
```

For the RISK_RATED_AREA_B and RISK_RATED_AREA_C, we delete their missing values.
```{r}
df<-df %>% drop_na(RISK_RATED_AREA_B)
df<-df %>% drop_na(RISK_RATED_AREA_C)
```

Let column PAYMENT_FREQUENCY NA=0
```{r}
df$PAYMENT_FREQUENCY[is.na(df$PAYMENT_FREQUENCY)] <- 0
df$PAYMENT_FREQUENCY<-as.factor(df$PAYMENT_FREQUENCY)
```

Calculate the client age
```{r}
library(lubridate)
birthday_year <- year(as.POSIXlt(df$P1_DOB, format="%d/%m/%Y"))
cover_year <- year(as.POSIXlt(df$COVER_START, format="%d/%m/%Y"))
df$client_age <- cover_year-birthday_year
```

Calculate the policy duration
```{r}
na_counts <- colSums(is.na(df))
print(na_counts)
```
```{r}
unique(df$POL_STATUS)
```
Delete "Unknown" observation in POL_STATUS
```{r}
df <- df[df$POL_STATUS != "Unknown", ]
```

To calculate the policy duration, first we delete the observation whose POL_STATUS is not "Live" and MTA_DATE is NA
```{r}
df$POL_STATUS<-as.character(df$POL_STATUS)
df <- subset(df, !(POL_STATUS != "Live" & is.na(MTA_DATE)))
```

Calculate the duration
```{r}
df$duration <- NA
cover_start_date <- as.Date(df$COVER_START, format = "%d/%m/%Y")

for (i in 1:nrow(df)) {
  # If status is Live
  if (df$POL_STATUS[i] == "Live") {
    # duration=2012/7/1- cover start date
    fixed_end_date <- as.Date("1/7/2012", format = "%d/%m/%Y")
    df$duration[i] <- (as.numeric(fixed_end_date - cover_start_date[i])) / 365
  } else {
    # If status is not Live, duration=MTA_date - cover start date
    MTA_date <- as.Date(df$MTA_DATE[i], format = "%d/%m/%Y")
    df$duration[i] <- (as.numeric(MTA_date - cover_start_date[i]))/365
  }
}

df <- df[df$duration != 0, ]
```

Delete time-related variable
```{r}
columns_to_exclude <- c("MTA_DATE", "COVER_START", "P1_DOB")
df <- df[, !names(df) %in% columns_to_exclude]
```

Let column PAYMENT_FREQUENCY NA=0
```{r}
df$PAYMENT_FREQUENCY[is.na(df$PAYMENT_FREQUENCY)] <- 0
df$PAYMENT_FREQUENCY<-as.factor(df$PAYMENT_FREQUENCY)
```

Finally we change the type of character variables so that it is easier for us to do EDA.
```{r}
columns_to_exclude <- c("MTA_DATE", "COVER_START", "P1_DOB")
df <- df[, !names(df) %in% columns_to_exclude]

char_vars <- names(df)[sapply(df, is.character)]
unique_list <- lapply(df[char_vars], unique)
unique_list<-head(unique_list, -1)
unique_list
```
Delete character variables that only have single value(eg: all "Y" and zero "N")
```{r}
one_value_subsets <- Filter(function(x) length(x) == 1, unique_list)
one_value_subsets
columns_to_exclude <- c("MTA_FLAG", "BUILDINGS_COVER",   "CONTENTS_COVER","AD_CONTENTS","AD_BUILDINGS","MTA_FLAG")
df <- df[, !names(df) %in% columns_to_exclude]
```

```{r}
# get variable names that are character type and have exactly two unique values
vars <- names(df)[sapply(df, function(x) is.character(x) & length(unique(x)) == 2)]

# convert selected variables to factor type
df[vars] <- lapply(df[vars], as.factor)
```


```{r}
df$Resiliated[df$POL_STATUS == 'Lapsed'] <- 1
df$Resiliated[df$POL_STATUS != 'Lapsed'] <- 0
columns_to_exclude <- c("POL_STATUS")
df <- df[, !names(df) %in% columns_to_exclude]
str(df)
```


```{r}
df$emp_status <- ifelse(df$P1_EMP_STATUS %in% c("E", "R", "S"), "Employed",
                        ifelse(df$P1_EMP_STATUS %in% c("U", "N", "A"), "Unemployed",
                               ifelse(df$P1_EMP_STATUS %in% c("H", "F", "V", "I", "C"), "Other", "Unknown")))

df$emp_status <- factor(df$emp_status, levels = c("Employed", "Unemployed", "Other"))


df$employment_status <- ifelse(df$P1_EMP_STATUS %in% c("E", "R", "S"), "Employed",
                               ifelse(df$P1_EMP_STATUS %in% c("U", "N", "A"), "Unemployed",
                                      "Other"))

df$employment_status <- factor(df$employment_status, levels = c("Employed", "Unemployed", "Other"))



df$OCC_STATUS <- factor(df$OCC_STATUS, levels = c("PH", "UN", "HH"))
df$PAYMENT_METHOD <- factor(df$PAYMENT_METHOD, levels = c("PureDD", "NonDD", "DD-Other"))
df$OCC_STATUS <- factor(df$OCC_STATUS, levels = c("PH", "UN", "HH"))
df$PAYMENT_METHOD <- factor(df$PAYMENT_METHOD, levels = c("PureDD", "NonDD", "DD-Other"))
```


EDA
```{r}
cor_matrix <- cor(df[, sapply(df, is.numeric)])

# 计算与MTA_APRP的相关系数
cor_with_MTA_APRP <- cor_matrix[, "MTA_APRP"]

# 绘制相关系数图
plot_cor_with_MTA_APRP <- ggplot(data = data.frame(x = names(cor_with_MTA_APRP), y = cor_with_MTA_APRP),
                                aes(x = x, y = y)) +
  geom_bar(stat = "identity", fill = "skyblue", width = 0.5) +
  labs(title = "Correlation with MTA_APRP", x = "Variable", y = "Correlation")

print(plot_cor_with_MTA_APRP)

```


