---
title: "432 data clean"
output: pdf_document
date: "2023-05-07"
---

```{r}
df <- read.csv("home_insurance.csv", header=T, na.strings=c("","NA"))
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

POL_STATUS is the response variable of our classification problem. We first delete all missing POL_STATUS.

```{r}
na_counts <- colSums(is.na(df))
print(na_counts)
na_cols <- names(na_counts[na_counts > 30000])
na_cols
```
Delete columns with too many missing values and inmeaningful
```{r}
columns_to_exclude <- c("P1_PT_EMP_STATUS", "CLERICAL", "CAMPAIGN_DESC", "UNSPEC_HRP_PREM", "Police", "P1_SEX", "SUBSIDENCE", "HP1_ADDON_PRE_REN", "HP1_ADDON_POST_REN", "HP2_ADDON_PRE_REN", "HP2_ADDON_POST_REN", "HP3_ADDON_PRE_REN", "HP3_ADDON_POST_REN", "GARDEN_ADDON_PRE_REN", "GARDEN_ADDON_POST_REN", "i")
df <- df[, !names(df) %in% columns_to_exclude]
```
POL_STATUS is the response variable of our classification problem. Our response variable MTA_APRP for regression problem. We drop all the missing value of these two variables.
```{r}
df<-df %>% drop_na(POL_STATUS)
df<-df %>% drop_na(MTA_APRP)
```

For the RISK_RATED_AREA_B and RISK_RATED_AREA_C, we delete their missing values.
```{r}
df<-df %>% drop_na(RISK_RATED_AREA_B)
df<-df %>% drop_na(RISK_RATED_AREA_C)
```

Let column PAYMENT_FREQUENCY NA=0
```{r}
df$PAYMENT_FREQUENCY[is.na(df$PAYMENT_FREQUENCY)] <- 0
df$PAYMENT_FREQUENCY<-as.factor(df$PAYMENT_FREQUENCY)
```

Calculate the client age
```{r}
library(lubridate)
birthday_year <- year(as.POSIXlt(df$P1_DOB, format="%d/%m/%Y"))
cover_year <- year(as.POSIXlt(df$COVER_START, format="%d/%m/%Y"))
df$client_age <- cover_year-birthday_year
```
Calculate the policy duration
```{r}

```






```{r}
str(df)
```

Let column PAYMENT_FREQUENCY NA=0
```{r}
df5$PAYMENT_FREQUENCY[is.na(df5$PAYMENT_FREQUENCY)] <- 0
df5$PAYMENT_FREQUENCY<-as.factor(df5$PAYMENT_FREQUENCY)
```

Finally we change the type of some variables so that it is easier for us to do EDA.
```{r}
df5$SUM_INSURED_BUILDINGS <-as.factor(df5$SUM_INSURED_BUILDINGS)
df5$LISTED <-as.factor(df5$LISTED)
df5$PAYING_GUESTS <-as.factor(df5$PAYING_GUESTS)
```



