---
title: "432 project"
output: html_document
date: "2023-03-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
df <- read.csv("home_insurance.csv", header=T, na.strings=c("","NA"))
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

POL_STATUS is the response variable of our classification problem. We first delete all missing POL_STATUS.

```{r}
sum(is.na(df$POL_STATUS))
```
```{r}
df2<-df %>% drop_na(POL_STATUS)
```


```{r}
unique(df2$POL_STATUS)
```
Delete unknown observation
```{r}
df3 <- df2[df2$POL_STATUS != "Unknown", ]
```

For regression problem
```{r}
summary(df3$MTA_APRP)
```
```{r}
sum(is.na(df3$MTA_APRP))
```

Delete the missing value of our response variable MTA_APRP for regression problem
```{r}
install.packages("data.table")
library(data.table)
df4<-data.table(df3 %>% drop_na(MTA_APRP))
```


Delete columns with too many missing values
```{r}
df5<-df4[,-c("P1_PT_EMP_STATUS","CLERICAL","CAMPAIGN_DESC")]
```

Let column PAYMENT_FREQUENCY NA=0
```{r}
df5$PAYMENT_FREQUENCY[is.na(df5$PAYMENT_FREQUENCY)] <- 0
df5$PAYMENT_FREQUENCY<-as.factor(df5$PAYMENT_FREQUENCY)
```

At that time, only QUOTE_DATE, MTA_DATE, RISK_RATED_AREA_B, RISK_RATED_AREA_C have missing values.

For RISK_RATED_AREA_B and RISK_RATED_AREA_C, both of which has small missing values:
```{r}
#replace NA with median
#Risk rate area B
mean_B <- median(df5$RISK_RATED_AREA_B, na.rm=TRUE)
df5$RISK_RATED_AREA_B <- replace_na(df5$RISK_RATED_AREA_B,mean_B)
#Risk rate area C
mean_C <- median(df5$RISK_RATED_AREA_C, na.rm=TRUE)
df5$RISK_RATED_AREA_C <- replace_na(df5$RISK_RATED_AREA_C,mean_C)
```

Since QUOTE_DATE and  MTA_DATE have a large number of missing values, we will not include these two variables in our models.

Finally we change the type of some variables so that it is easier for us to do EDA.
```{r}
df5$SUM_INSURED_BUILDINGS <-as.factor(df5$SUM_INSURED_BUILDINGS)
df5$LISTED <-as.factor(df5$LISTED)
df5$PAYING_GUESTS <-as.factor(df5$PAYING_GUESTS)
```

Save the origin data
```{r}

```


EDA part

```{r}
df5=data.table(df5)
total <- nrow(df5)
status_groups <- df5[, .(count=.N, percent=round((.N/total)*100, 2)), by = POL_STATUS]
status_groups <- status_groups[order(count, decreasing = TRUE)]
status_groups
pieLabels <- paste(status_groups$POL_STATUS,' ', status_groups$percent, '%')
library(plotrix)
install.packages("plotrix", version = "3.7-7")
pie(status_groups$count,labels=pieLabels,explode=0.1, radius=0.8,height=0.1, col=rainbow(length(status_groups$POL_STATUS)),
  	main="Pie Chart of Policy Status ")
```



```{r}
status_groups[POL_STATUS != 'Lapsed', POL_STATUS:= "Non Resiliated"]
status_groups[POL_STATUS == 'Lapsed', POL_STATUS:= "Resiliated"]
status_groups <- status_groups[, .(count=sum(count), percent = round((.N*100)/sum(count), 2)), by = POL_STATUS]
status_groups[,percent := round((count*100)/sum(count), 2)]
status_groups <- status_groups[order(count, decreasing = TRUE)]
status_groups
pieLabels <- paste(status_groups$POL_STATUS,' ', status_groups$percent, '%')
pie <- pie(status_groups$percent,labels=pieLabels,explode=0.1, radius=0.8,height=0.1, col=rainbow(length(status_groups$POL_STATUS)),
  	main="Pie Chart of Resiliation")
```

```{r}
df5$Resiliated[df5$POL_STATUS == 'Lapsed'] <- 1
df5$Resiliated[df5$POL_STATUS != 'Lapsed'] <- 0
table(df5$Resiliated)
```
```{r}
df6=df5
```


Find the employment ststus of each policy holder
```{r}
status_client <- df5[!is.na(P1_EMP_STATUS), .(count=.N), by = P1_EMP_STATUS]
status_client <- status_client[order(count, decreasing = TRUE)]
status_client$percent <- status_client$count / sum(status_client$count) * 100
```

R represents retired. E represents employed. S represents student. H represents house person.U represents unemployed. N represents not available.
There are about 11 employment status. The Retired people form the overwhelmingly majority. The Employees and Students come at a very distant second and third respectively.
```{r}
ggplot(data = status_client, aes(x = P1_EMP_STATUS, y = count)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  geom_text(aes(label = paste0(round(percent, 1), "%"), y = count), vjust = -0.5) +
  theme_minimal() +
  labs(x = "Employment Status", y = "Count", title = "Count of Clients by Employment Status")
```
R represents retired. E represents employed. S represents student. H represents house person.U represents unemployed. N represents not available.
There are about 11 employment status. The Retired people form the overwhelmingly majority(76.5%). The Employees(19.5%) and Students(2.1%) come at second and third respectively.

```{r}
month_order <- c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec')
day_order <- c('Mon', 'Tue', 'Wed', 'Thu', 'Fri', 'Sat', 'Sun')
df5$quotemonth_n <- month(as.POSIXlt(df5$QUOTE_DATE, format="%m/%d/%Y"))
df5$covermonth_n <- month(as.POSIXlt(df5$COVER_START, format="%d/%m/%Y"))
#Name the columns
quotesmonthDF <- data.frame(month_n = df5$quotemonth_n )
coversmonthDF <- data.frame(month_n = df5$covermonth_n)
#head(quotesmonthDF, 2)
#Avoid the null values and make the group by each month to get the monthly total
quotesmonthgroup <- data.table(quotesmonthDF)
quotesmonthgroup <-quotesmonthgroup[month_n <= 12]
quotesmonthgroup <-quotesmonthgroup[(order(month_n)), .(count=.N), by=month_n]

coversmonthgroup <- data.table(coversmonthDF)
coversmonthgroup <-coversmonthgroup[month_n <= 12]
coversmonthgroup <-coversmonthgroup[(order(month_n)), .(count=.N), by=month_n]
#Add the name of the month

quotesmonthgroup$month_s <- month_order[quotesmonthgroup$month_n ]
coversmonthgroup$month_s <- month_order[coversmonthgroup$month_n ]
head(quotesmonthgroup, 12)
head(coversmonthgroup, 12)



df5$quotemonth_s <- month_order[df5$quotemonth_n ]
df5$covermonth_s <- month_order[df5$covermonth_n ]

barplot(quotesmonthgroup)
barplot(coversmonthgroup)
```






